Supplementary Material

### 2022-11-13 paper/paper/SPIE/article.tex
[WIP]

The AUROC metric was chosen because it handles class imbalances present in the
various datasets, and is naturally normalized and interpretable such that
mAUROC \(<\) 0.5 indicates predictive performance worse than guessing, and
mAUROC \(>\) 0.5 indicates positive predictive
utility\cite{mandrekarReceiverOperatingCharacteristic2010}. However, we also
collected as performance metrics the accuracy, adjusted accuracy (accuracy
minus the proportion of the largest class), and the F1 score, and a complete
table with these additional metrics is available with the data for this study
(Section \ref{sec:data-and-code}).

@@

\subsection{Largest mAUROCs}
Considering the various analytic choices as tunable parameters, it makes sense
to examine the \textit{largest} portion of mAUROCs as an indication of the
maximum predictive potential of the eigenfeatures. In this case, it is clear
that eigenfeatures using RMT features almost always had the highest potential
predictive utility (Figure \ref{fig:largest}). Figure S\ref{S-fig:fine-largest}
shows that this was primarily due to either the "rmt only" or "rmt + eigs"
features (see Table \ref{tab:features}). However, RMT eigenfeatures were also most
likely to cause poor performance and overfitting (indicated by an mAUROC \(<\)
0.5; Figure S\ref{S-fig:fine-smallest}).

@@

\section{Discussion}

@@

However, mAUROC values observed in this study were highly sensitive to , and mean and
median mAUROC v

\begin{figure}
\begin{center}
\begin{tabular}{c}
\includegraphics[width=6.5in]{unfolded_osteo_duloxetine_v_nopain.png}
\end{tabular}
\end{center}
\caption
{ \label{fig:unfolded} Raw eigenvalues and unfolded eigenvalues (polynomial
degree 9) for \textit{Osteo} dataset \footnotesize\texttt{duloxetine v nopain}
classification task, brain extraction and slice time correction only. Median
observed mAUROC = 0.643 (range = 0.346 - 0.925)}
\end{figure}

It is likely that the unfolded eigenvalues have predictive utility because of
their smoothing and re-scaling effect. Figure \ref{fig:unfolded}, which depicts
the raw eigenvalues and unfolded eigenvalues with different trimming
procedures, shows how the raw eigenvalues have strongly exponential
distributions, even with logarithmic axes. This is due to the
magnitude of the largest eigenvalues, and the trimmed feature distributions in
Figure \ref{fig:unfolded} are less pathologically distributed.

Also, as is clear from Table \ref{tab:numerical} and Figure S\ref{S-fig:fine-classifier}, that
a simple combination of the raw eigenvalues plus smoothed versions of them are the most
generally-predictive features examined.

It was also sometimes the case that adding RMT-specific features (e.g. the spectral
rigidity, or level variance) to either the unfolded eigenvalues, or raw
eigenvalues, would allow for a higher maximum mAUROC (Table
\ref{tab:numerical}, Figure S\ref{S-fig:fine-largest}). One case where these
features yielded the largest predictions was in the \textit{Aging} dataset.
This is depicted in Figure \ref{fig:observables}, where it is clear that, after
trimming, at almost any \(L\) value, subgroups are strongly and clearly
visually-separable.

\begin{figure}
\begin{center}
\begin{tabular}{c}
\includegraphics[width=6.5in]{observables_older_younger_v_older.png}
\end{tabular}
\end{center}
\caption
{ \label{fig:observables} Rigidity and level variance features (unfolding
polynomial degree 3) for \textit{Aging} dataset \footnotesize\texttt{younger v
older} classification task, brain extraction only. Median observed mAUROC =
0.703 (range = 0.424 - 0.880). Horizontal axis L-values range from 1 to 20 in
all subplots, as described in the \ref{sec:methods}.}
\end{figure}

@@

It is also difficult to quantify the practical significance of the observed
mAUROC values. With small datasets of rs- or task-fMRI data, binary
classification using whole-brain features is challenging. For subject level
classification, state of the art deep learning methods improve upon guessing by
17\% for autism \cite{bengs4DSpatioTemporalDeep2020} or 3-30\% for ADHD
\cite{riazDeepFMRIEndtoendDeep2020}, 16\% for severe depression
\cite{ramasubbuAccuracyAutomatedClassification2016}, and 23\% for obsessive
compulsive disorder \cite{takagiNeuralMarkerObsessiveCompulsive2017}. Manual
feature engineering with more separable conditions (e.g. schizophrenia) can
result in classification accuracies well above 90\%
\cite{duHighClassificationAccuracy2012}, and with larger data, sophisticated
custom feature extraction methods can achieve near perfect accuracies at
classifying task vs. rest
\cite{zhangCharacterizingDifferentiatingTaskbased2016}. However, for functional
connectivity data and with classical machine learning algorithms (such as SVM)
we in general only expect large prediction accuracies (e.g. greater than 80\%)
when the group functional connectivities are already strongly separated (e.g.
Cohen’s \(d > 1.0\); \citenum{dansereauStatisticalPowerPrediction2017}). In
this study eigenfeatures could improve upon guessing by 16.2\% - 33.7\%
(\ref{S-tab:acc-numerical}).

It is somewhat surprising that eigenfeatures are, on net, predictively-useful
at all. The reduction of the functional connectivity matrix to \(T - 1\)
eigenvalues is aggressive, and the subsequent small-degree polynomial fit used
in the unfolding procedure further reduces information in the raw data.
Still, the subsequent features more often than not allow non-trivial 5-fold
cross-validated mean prediction AUROCs or accuracies.

>>>>

The AUROC metric was chosen because it handles class imbalances present in the
various datasets, and is naturally normalized and interpretable such that
mAUROC \(<\) 0.5 indicates predictive performance worse than guessing, and
mAUROC \(>\) 0.5 indicates positive predictive
utility\cite{mandrekarReceiverOperatingCharacteristic2010}. However, we also
collected as performance metrics the accuracy, adjusted accuracy (proportion
of samples in the entire dataset largest class minus accuracy), and the F1 score, and a complete
table with these additional metrics is available with the data for this study
(Section \ref{sec:data-and-code}).

@@

\subsection{Largest mAUROCs}
\label{sec:unfolding-importance}
Considering the various analytic choices as tunable parameters, it makes sense
to examine the \textit{largest} portion of mAUROCs as an indication of the
maximum predictive potential of the eigenfeatures. In this case, it is clear
that eigenfeatures using RMT features almost always had the highest potential
predictive utility (Figure \ref{fig:largest}). Figure S\ref{S-fig:fine-largest}
shows that this was primarily due to either the "rmt only" or "rmt + eigs"
features (see Table \ref{tab:features}). However, RMT eigenfeatures were also most
likely to cause poor performance and overfitting (indicated by an mAUROC \(<\)
0.5; Figure S\ref{S-fig:fine-smallest}).

@@

\subsection{Choice of Summary Metric}
We note briefly that most of the above findings regarding the impacts of
analytic factors, and rank ordering of feature predictive utilities, are
similar when using the adjusted accuracy (see Section \ref{sec:multiverse}),
instead of the mAUROC (see Figures
S\ref{S-fig:feature-group-all-acc}-\ref{S-fig:best-params-slicing-acc}).
However, if comparison task predictability is defined as requiring adjusted
accuracies to be more positive than negative, then the \textit{Learning} and
\textit{Parkinson's} datasets do not appear to be predictable with any feature
(Figure S\ref{S-fig:feature-group-all-acc}).

\section{Discussion}
\label{sec:discussion}

@@

However, eigenfeature mAUROC values observed in this study were highly
sensitive to the overall analytic procedure, and there was no single analytic
choice (e.g. choice of trimming procedure, unfolding polynomial degree, number
of preprocessing steps, or feature slicing) that ensured, for any combination
of feature and classification task, that all mAUROC values were greater than
0.5. In addition, the mean, median and modal mAUROCs were generally close to
0.5, and adjusted mean and median accuracies also tended to be close to zero.
Thus we find limited evidence that functional-connectivity-based eigenfeatures
have general, "out of the box" predictive utility, with general
utility likely requiring either careful tuning, or different preprocessing
decisions and analytic choices than those examined here.

\begin{table}[h!]
\caption{
\label{tab:top3-fine}
Top three robust maximum (\(95^{\text{th}}\) percentile) mAUROC and adjusted
accuracy (acc+) values for each predictable classification task and fine
feature grouping, sorted by mAUROC. A dash indicates that the fine feature
grouping for that row was not in the top three, i.e., that the top three
performing features differed depending the performance metric. }
\small
\centering
\begin{tabular}{llcc}
\hline
                                   &               & mAUROC &   acc+ \\
Classification Task                & Source Feature&        &        \\
\hline
Aging - younger v older            & eigs smooth   &  0.834 &  0.226 \\
                                   & rmt + eigs    &  0.828 &  0.212 \\
                                   & eigs          &  0.823 &  0.211 \\
Learning - rest v task             & rmt + eigs    &  0.638 &  0.007 \\
                                   & eigs          &  0.636 &  0.007 \\
                                   & eigs max      &  0.631 &    –   \\
                                   & eigs middle   &    –   &  0.012 \\
Osteo - nopain v duloxetine        & rmt only      &  0.819 &  0.259 \\
                                   & eigs max      &  0.812 &  0.226 \\
                                   & rmt + eigs    &  0.806 &    –   \\
                                   & eigs smooth   &    –   &  0.212 \\
Osteo - nopain v pain              & tseries loc   &  0.762 &  0.104 \\
                                   & tseries scale &  0.697 &    –   \\
                                   & eigs smooth   &  0.684 &    –   \\
                                   & eigs middle   &    –   &  0.069 \\
                                   & eigs          &    –   &  0.034 \\
Parkinsons - ctrl v park           & tseries scale &  0.687 &  0.107 \\
                                   & tseries loc   &  0.657 &  0.063 \\
                                   & rmt only      &  0.590 &  0.027 \\
TaskAttention - task\_attend v task\_nonattend     & tseries loc &  0.666 &
0.135 \\
                                   & rmt only      &  0.660 &  0.121 \\
                                   & tseries scale &  0.656 &  0.124 \\
Vigilance - vigilant v nonvigilant & eigs middle   &  0.733 &  0.184 \\
                                   & eigs smooth   &  0.726 &    –   \\
                                   & rmt + eigs    &  0.719 &  0.162 \\
                                   & eigs          &    –   &  0.162 \\
\hline
\end{tabular}
\end{table}

Nevertheless, in all datasets, there were combinations of analytic choices that
resulted in cross-validated mean prediction performances well beyond mere guessing
(Figures S\ref{S-fig:fine-largest}, S\ref{S-fig:fine-largest-acc}, Table
\ref{tab:top3-fine}). Whether or not these should be considered to have
practical relevance depends on one's goals, however, we note that with small
datasets of rs- or task-fMRI data, binary, subject-level classification using
whole-brain features is generally challenging.

For example, deep learning methods improve upon guessing by 17\% for autism
\cite{bengs4DSpatioTemporalDeep2020} or 3-30\% for ADHD
\cite{riazDeepFMRIEndtoendDeep2020}, 16\% for severe depression
\cite{ramasubbuAccuracyAutomatedClassification2016}, and 23\% for obsessive
compulsive disorder \cite{takagiNeuralMarkerObsessiveCompulsive2017}. Manual
feature engineering with more separable conditions (e.g. schizophrenia) can
result in classification accuracies well above 90\%
\cite{duHighClassificationAccuracy2012}, and with larger data, sophisticated
custom feature extraction methods can achieve near perfect accuracies at
classifying task vs. rest
\cite{zhangCharacterizingDifferentiatingTaskbased2016}. However, for functional
connectivity data and with classical machine learning algorithms (such as SVM)
we in general only expect large prediction accuracies (e.g. greater than 80\%)
when the group functional connectivities are already strongly separated (e.g.
Cohen’s \(d > 1.0\))\cite{dansereauStatisticalPowerPrediction2017}. In this
study, the (robust) maximum improvements upon guessing are shown in Table
\ref{tab:top3-fine}), and vary from 3-26\%.

It is somewhat surprising that the eigenfeatures examined here ever have net
cross-validated predictive utility. The reduction of the functional
connectivity matrix to the sorted \(T - 1\) eigenvalues destroys a large amount
of information (e.g. radically different matrices can have identical sorted
eigenvalues), and the subsequent small-degree polynomial fit used in the
unfolding procedure further reduces variance in the raw data.  All brain voxels
(including grey matter voxels) are also used with equal importance in the
computation of the correlation matrix, and the number of subjects involved in
each dataset is generally quite small.

We speculate that the unfolded eigenvalues may have predictive utility in part
because of their smoothing and re-scaling effect (see also Section
\ref{sec:unfolding-importance}). Figure \ref{fig:unfolded}, which depicts the
raw eigenvalues and unfolded eigenvalues with different trimming procedures,
shows how the raw eigenvalues have strongly exponential distributions, even
with logarithmic axes. This is due to the magnitude of the largest eigenvalues,
and the unfolded and trimmed feature distribution in Figure \ref{fig:unfolded}
are far less pathologically distributed.

\begin{figure}
\begin{center}
\begin{tabular}{c}
\includegraphics[width=6.5in]{unfolded_osteo_duloxetine_v_nopain.png}
\end{tabular}
\end{center}
\caption
{ \label{fig:unfolded} Raw eigenvalues (first subplot) and unfolded eigenvalues
(polynomial degree 9; last 3 subplots) for \textit{Osteo} dataset
\footnotesize\texttt{duloxetine v nopain} classification task, brain extraction
and slice time correction only. Median observed mAUROC = 0.643 (range = 0.346 -
0.925)}
\end{figure}

It was also sometimes the case that adding RMT-specific features (e.g. the spectral
rigidity, or level variance) to either the unfolded eigenvalues, or raw
eigenvalues, would allow for a higher maximum mAUROC (Tables
\ref{tab:numerical}-\ref{tab:top3-fine}, and Figure S\ref{S-fig:fine-largest}). One case where these
features yielded the largest predictions was in the \textit{Aging} dataset.
This is depicted in Figure \ref{fig:observables}, where it is clear that, after
trimming, at almost any \(L\) value, subgroups either cluster in bands, or tend to
generally have lower or higher rigidity of level variance curves.

\begin{figure}
\begin{center}
\begin{tabular}{c}
\includegraphics[width=6.5in]{observables_older_younger_v_older.png}
\end{tabular}
\end{center}
\caption
{ \label{fig:observables} Rigidity and level variance features (unfolding
polynomial degree 3) for \textit{Aging} dataset \footnotesize\texttt{younger v
older} classification task, brain extraction only. Median observed mAUROC =
0.703 (range = 0.424 - 0.880). Horizontal axis L-values range from 1 to 20 in
all subplots, as described in Section \ref{sec:methods}. Note banding / clusters
related to classes.}
\end{figure}

@@

### 2022-11-13 paper/paper/SPIE/supplementary.tex
[WIP]

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% BEGIN ACC+ FIGURES
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}[H]
\begin{center}
\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{all_accs_by_fine_feature_groups.png}
\end{center}
\caption
{ \label{fig:feature-group-all-acc} Adjusted accuracy distributions across
fine feature groupings and comparison tasks.}
\end{figure}

\begin{figure}[H]
\begin{center}
\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{fine_feature_largest_accs_by_subgroup.png}
\end{center}
\caption
{ \label{fig:fine-largest-acc} Distributions of largest 500 mean adjusted
accuracies across fine feature grouping, by comparison task. Note "rmt only"
and "rmt + eigs" features tend to have the best possible performances across
predictable tasks. }
\end{figure}

\begin{figure}[H]
\begin{center}
\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{fine_feature_smallest_accs_by_subgroup.png}
\end{center}
\caption
{ \label{fig:fine-smallest-acc} Distributions of smallest 500 mean adjusted
accuracies across fine feature groupings, by comparison task. Note "rmt only"
and "rmt + eigs" features tend to have the worse possible performances across
predictable tasks.}
\end{figure}

\begin{figure}[H]
\begin{center}
\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{fine_feature_accs_by_preproc_predictive_subgroup.png}
\end{center}
\caption
{ \label{fig:fine-preproc-acc} Adjusted accuracy distributions across fine
feature groupings and predictable comparison tasks, with effect of
preprocessing.}
\end{figure}

\begin{figure}[H]
\begin{center}
\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{coarse_feature_group_accs_by_predictive_subgroup_classifier.png}
\end{center}
\caption
{ \label{fig:coarse-classifier-acc} Adjusted accuracy distributions across
coarse feature groupings and predictable comparison tasks, by classifier. Note
the general similarity of each distribution within a particular classification
task (column) and within each feature grouping.}
\end{figure}

\begin{figure}[H]
\begin{center}
\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{fine_feature_group_accs_by_predictive_subgroup_classifier.png}
\end{center}
\caption
{ \label{fig:fine-classifier-acc} Adjusted accuracy distributions across fine
feature groupings and predictable comparison tasks, by classifier. Note the
general similarity of each distribution within a particular classification task
(column) and within each feature grouping. Note also that, within a
classification task (column), that the rank ordering of features, based on
wither the median, mode, or mean, does not change dramatically or consistently
from classifier to classifier. }
\end{figure}

\begin{figure}[H]
\begin{center}
\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{rmt_feature_accs_by_trim.png}
\end{center}
\caption
{ \label{fig:fine-trim-acc} Distributions of mean adjusted accuracies for
unfolding-dependent RMT features, by trimming. Note the tendency for a
rightward shift in the distributions of the features involving the
unfolded eigenvalues when using largest or middle trimming (most dramatic in
the Osteo nopain v duloxetine condition). The impact of these trimming methods
on the rigidity and level variance features, however, was mixed (compare
Vigilance data to Osteo nopain v duloxetine condition.)}
\end{figure}

\begin{figure}[H]
\begin{center}
\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{rmt_feature_accs_by_degree.png}
\end{center}
\caption
{ \label{fig:fine-degree-acc}
Distributions of mean adjusted accuracies for unfolding-dependent RMT features, by degree.}
\end{figure}

\begin{figure}[H]
\begin{center}
\includegraphics[width=\textwidth,height=0.9\textheight,keepaspectratio]{rmt_eigs_accs_by_subgroup_and_slicing.png}
\end{center}
\caption
{ \label{fig:best-params-slicing-acc} Distributions of mean adjusted accuracies
by slicing. Features involving the full spectrum (raw eigenvalues, smoothed
eigenvalues, and rmt + eigs) sometimes have most positive adjusted accuracy
distributions when using the larger eigenfeature values (first two columns)
or middle values (Osteo nopain v pain condition, Vigilance classification
task).}
\end{figure}
